### 2025-09-20 Doctor_Paper_Modification/chapters/publications.tex

% !TeX root = ../main.tex

\begin{publications}

\section*{已发表论文}

\begin{enumerate}
\item XU X H, XIE J L, ZHANG L F, et al. Improvement of transmittance using groove structured surface for microwave imaging diagnostics in tokamak plasmas [J]. Int Conf Infra Milli, 2020.
\item HAN D Q, XIE J L, HUSSAIN A, Xu X H,et al. In situ relative self-dependent calibration of electron cyclotron emission imaging via shape matching [J]. Review of Scientific Instruments, 2018, 89(10).
\item GAO B X, XIE J L, ZHAO Z L, Xu X H,et al. Diagnostic capacity of electron cyclotron emission imaging system with continuous large observation area on EAST tokamak [J]. Review of Scientific Instruments, 2018, 89(9).
\item
LIAO W, GAO F, XU X, et al. Bench test of microwave imaging reflectometry system for EAST tokamak [J]. Journal of Instrumentation, 2020, 15(03): C03036.
\item
高飞雪, 刘贤子, 渠承明,徐新航, et al. EAST 装置微波反射成像系统光学性能评测 [J]. 核聚变与等离子体物理, 2022, 42(2): 187.
\end{enumerate}

\section*{待发表论文}

\begin{enumerate}
\item Xu Xinhang,Xie Jinlin ,et al.Constraining Electron Energy in static electric Field via the Anomalous Doppler Resonant with External Electromagnetic Waves.
\item Xu Xinhang,Xie Jinlin ,et al.Trapping analysis of a magnetic electron by a circularly polarized electromagnetic wave in static electric field.
\end{enumerate}

%\section*{研究报告}
%\begin{enumerate}
%\item A A A A A A A A A
%\item A A A A A A A A A
%\item A A A A A A A A A
%\end{enumerate}

\end{publications}