### 2025-10-09 Doctor_Paper_Modification/待签字文件/combine_factory/pdf_combine.tex
presentation

% !TeX encoding = UTF-8
% !TeX program = xelatex
% !TeX spellcheck = en_US

\documentclass{article} % You need a document class

\usepackage[margin=1in]{geometry} % optional, sets margins
\usepackage[justification=centering]{caption}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{pdfpages} % required for including PDF pages

\setcounter{secnumdepth}{2}

\begin{document}

% Include the first PDF (first page only)
\includepdf[pages=1,fitpaper=true]{groove_paper.pdf}

% Include the second PDF (pages 2 to 9)
\includepdf[pages=2,fitpaper=true]{Improvement_of_transmittance_using_groove_structured_surface_for_microwave_imaging_diagnostics_in_tokamak_plasmas.pdf}

\end{document}